Finite model property for residuated semigroups and related remarks

### 2023-03-12 residuated-semigroups/Text.tex
Please enter the commit message for your changes. Lines starting
consider two particular examples

\item {\bf Left redisual move}
\item {\bf Right residual move}
\end{enumerate}
\end{enumerate}

@@

\begin{enumerate}
\item $\exists$ has a winning stragery in $\mathcal{G}_{\omega}(\mathcal{A})$
\item $\mathcal{A}$ is representable.
\end{enumerate}
\end{theorem}

@@

\section{Games for the FMP}

>>>>

\item {\bf Left redisual move}

\item {\bf Right residual move}

\end{enumerate}
\end{enumerate}

@@

\begin{enumerate}
\item $\exists$ has a winning stragery in $\mathcal{G}_{\omega}(\mathcal{A})$

\begin{center}
$rep(a) = \{ (x,y) \: | \: l(x, y) \leq a \}$
\end{center}

\begin{enumerate}
\item $rep(a ; b) = rep(a) ; rep(b)$
\item $rep(a \setminus b) = rep(a) \setminus rep(b)$
\item $rep(a / b) = rep(a) / rep(b)$
\end{enumerate}

\item $\mathcal{A}$ is representable.
\end{enumerate}
\end{theorem}

@@

\section{Step-by-step construction for the FMP}

\subsection{Example 1}

Consider a sequent $\varphi_1 \bullet \varphi_2 \rightarrow \varphi_1 \bullet (\psi_1 \setminus \psi_2)$, which is known to unprovable in the Lambek calculus. We would like to construct a finite model on a network such that there exists $(u, v) \nvDash \varphi_1 \bullet \varphi_2 \rightarrow \varphi_1 \bullet (\psi_1 \setminus \psi_2)$, that is, $(u, v) \in ||\varphi_1|| ; ||\varphi_2||$ and $(u, v) \notin ||\varphi_1|| ; (||\psi_1|| \setminus ||\psi_2||)$.

The construction is step-by-step:
\begin{enumerate}
\item {\bf Step 0}

We construct the edge $(u, v)$ labelled by a pair of sets $\lambda_0$ and $\delta_0$:

\centerline{
\xymatrix{
u \ar[rr]^{\lambda_0}_{\delta_0} && v
}
}

\begin{center}
$\lambda_{0}(u, v) = \{ \varphi_1 \bullet \varphi_2, -(\varphi_1 \bullet (\psi_1 \setminus \psi_2)) \}$

$\delta_{0}(u, v) = \{ \varphi_1 \bullet \varphi_2\}$
\end{center}

\item {\bf Step 1}

We have to create a witness for $\varphi_1 \bullet \varphi_2 \in \delta_{0}(u, v)$ by adding a vertice $w$:

\centerline{
\xymatrix{
& w \ar[ddr]^{\lambda_1}_{\delta_1} \\
\\
u \ar[uur]^{\lambda_1}_{\delta_1}
\ar[rr]^{\lambda_0}_{\delta_0} && v
}
}

\begin{center}
$\lambda_1(u, w) = \{ \varphi_1 \}$

$\lambda_1(w, v) = \{ \varphi_2, -(\psi_1 \setminus \psi_2)\}$

$\delta_1(u, w) = \emptyset$

$\delta_1(w, v) = \{ - (\psi_1 \setminus \psi_2) \}$
\end{center}

\item {\bf Step 2}

We create a witness $w'$ for $- (\psi_1 \setminus \psi_2) \in \delta_1(w, v)$

\centerline{
\xymatrix{
w \ar[ddrr]^{\lambda_1}_{\delta_1} && w' \ar[ll]_{\lambda_2}^{\delta_2} \ar[dd]_{\lambda_2}^{\delta_2} \\
\\
u \ar[uu]^{\lambda_1}_{\delta_1}
\ar[rr]^{\lambda_0}_{\delta_0} && v
}
}

By putting:
\begin{center}
$\lambda_2(w', w) = \{ \psi_1 \}$

$\lambda_2(w', v) = \{ -\psi_2 \}$

$\delta_2(w', w) = \emptyset$

$\delta_2(w', v) = \emptyset$
\end{center}

\end{enumerate}

Now let us describe $||\varphi_1|| ; ||\varphi_2||$ as follows:

TODO:

\subsection{Example 2}

Let us consider a sequent $LC \not\vdash \varphi_1 \setminus (\varphi_2 \bullet \varphi_3) \rightarrow \varphi_1 \setminus \psi$

\begin{enumerate}
\item {\bf Step 0}

Introduce a wedge $(u, v)$ with labelling $\lambda_0$ and $\delta_0$

\centerline{
\xymatrix{
u \ar[rr]^{\lambda_0}_{\delta_0} && v
}
}

where

\begin{center}
$\lambda_0(u, v) = \{\varphi_1 \setminus (\varphi_2 \bullet \varphi_3), -(\varphi_1 \setminus \psi)\}$

$\delta_0(u, v) = \{ -(\varphi_1 \setminus \psi) \}$
\end{center}

\item {\bf Step 1}

We create a witness $w$ for $-(\varphi_1 \setminus \psi) \in \delta_0(u, v)$

\centerline{
\xymatrix{
& w \ar[ddr]^{\lambda_1}_{\delta_1} \ar[ddl]_{\lambda_1}^{\delta_1} \\
\\
u \ar[rr]^{\lambda_0}_{\delta_0} && v
}
}
where
\begin{center}
$\lambda_1(w, u) = \{ \varphi_1 \}$

$\lambda_1(w, v) = \{ -\psi, \varphi_1 \bullet (\varphi_1 \setminus (\varphi_2 \bullet \varphi_3)), \varphi_2 \bullet \varphi_3\}$

$\delta_1(w, u) = \emptyset$

$\delta_1(w, v) = \{ \varphi_2 \bullet \varphi_3 \}$

\end{center}

\item {\bf Step 2}

We have to create a witness $w'$ for $\varphi_2 \bullet \varphi_3 \in \delta_1(w, v)$

\centerline{
\xymatrix{
w \ar[ddrr]^{\lambda_1}_{\delta_1} \ar[dd]_{\lambda_1}^{\delta_1} \ar[rr]^{\lambda_2}_{\delta_2} && w' \ar[dd]_{\lambda_2}^{\delta_2} \\
\\
u \ar[rr]^{\lambda_0}_{\delta_0} && v
}
}

\begin{center}
$\lambda_2(w, w') = \{ \varphi_2 \}$

$\lambda_2(w', v) = \{ \varphi_3 \}$

$\delta_2(w, w') = \emptyset$

$\delta_2(w', v) = \emptyset$
\end{center}

\end{enumerate}

\subsection{General case}

Consider a sequent $\Gamma \rightarrow \varphi$ such that $LC \not\vdash \Gamma \rightarrow \varphi$. We have to construct a finite relational model over some network such that $(u, v) \not\models \Gamma \rightarrow \varphi$, that is, $(u, v) \models \Gamma$ and $(u, v) \not\models \varphi$.
